Rapid technological progress in white light-emitting diodes \\ and its sources in innovation and technology spillovers

### 2024-03-30 2_SSL_NE/article/article_supplementary.tex
edits in supplement

The historical development development of light-emitting diodes from a novelty semiconductor experiment into powerful lighting devices has received a considerable amount of attention following the recognition of pioneering work on blue LED by Japanese researchers Shuji Nakamura, Isamu Akasaki and Hiroshi Amano with the 2014 Nobel Prize in Physics \cite{Nakamura2015}. The sources and disaggregated contributions of subsequent improvements in LED technology, however, have not been documented in the literature in a systematic fashion, though there were notable publications reporting on the progress in the design of devices \cite{Shchekin2006,krames2007led,laubsch2009high,hahn2014development}, or the technological improvements  underlying the improvement in overall device efficiency for best performers in 2009 \cite{tsao2010solid} and 2016 \cite{pattison2017solid}.

@@

Due to these limitations, any recommendations on future improvements in LED technology based on these studies are fragmented at best. Furthermore, the effect of individual innovations and technology spillovers, which has been investigated in solar PV \cite{kavlak2018evaluating,kolesnikov2021spillovers,nemet2019solar}, and to some extent in lithium-ion batteries \cite{Stephan2021}, has not yet been studied in the context of lighting.

In our study, we overcome the limitations of previous literature by, first, synthesizing data provided in multiple publications across an ensemble of various cost, performance and consumer experience metrics for tracking progress in LED technology over an extended period of time. Second, our key advantage and contribution compared to these publications is a multi-method approach that we take to fill in the data gaps and augment information obtained from the literature sources by our own performance calculations, a bottom-up cost model and a series of expert interviews. This enabled us to synthesize evidence from multiple sources and, for the first time, reconstruct a comprehensive picture of the historical progress in white LED technology throughout its entire history.

@@

\begin{figure}[H]
\centering
\includegraphics[width=15.5cm]{./figures/subefficiencies_progress.pdf}
\caption{\textbf{Historical improvements in LED sub-efficiencies.} Panels A1-A2 show data for key metrics used to calculate forward voltage efficiency $\eta_{Vf}$ and droop efficiency $\eta_{droop}$. Panels B1-B2 show data for internal quantum efficiency $\eta_{IQ}$ and light extraction efficiency $\eta_{LE}$. Spectral data related to spectral sub-efficiency is shown in \cref{fig:phosphor_spectrum}. Additional context for the panels is provided below. Source: own synthesis of data from the list of sources provided in \cref{sec:sources}. For an illustration of the different chip architectures shown, compare \cref{sec:chip_architecture}. Data on average device performance adapted from U.S. Department of Energy (DOE) Reports \cite{doe_ssl_multiyear_2007}\cite{doe_ssl_multiyear_2008}\cite{doe_ssl_multiyear_2013}\cite{doe_ssl_rnd_2016}\cite{doe_ssl_rnd_2018} and an Osram company report \cite{beale_leds_2015}.}
\label{fgr:subeff}
\vspace{-20mm}
\end{figure}

@@

\begin{table}[H]
    \caption{\textbf{LED innovations and technology improvements affecting key device sub-efficiencies.} Note this table is continued in \cref{tab:innovations_2}.}
    \begin{NiceTabularX}{\textwidth}{ |l|X|X|l|l| }
        \hline
        \multicolumn{5}{c}{Forward Voltage Efficiency (FVE) $\eta_{Vf}$} \\
        \hline
            \textit{Year} & \textit{Innovation} & \textit{Area of Improvement} & \textit{Spillover} & \textit{Source} \\
        \hline
            1999 & Indium Tin Oxide \newline Current spreading layer & Contact resistance & Yes & \cite{margalith1999indium}\\
        \hline
            1998 & Digitated electrodes & Contact resistance & No & \cite{steigerwald2001electrode} \\
        \hline
            Ongoing & Epitaxy improvements \newline and better doping & Polarization and \newline bulk resistance & No & I \\
        \hline
            1998 & Silver p-Contacts & Contact resistance & No & \cite{kondoh2001nitride} \\
        \hline
        \multicolumn{5}{c}{Light extraction efficiency (LEE) $\eta_{LE}$} \\
        \hline
            \textit{Year} & \textit{Innovation} & \textit{Area of improvement} & \textit{Spillover} & \textit{Source} \\
        \hline
            $<$ 2003 & Optimization for \newline cavity effects & Reduces self-interference of quantum well & No & I, \cite{Shen2003} \\
        \hline
            1993 & Chip surface randomization & Total reflection and absorption & No & I, \cite{bergh1973surface}\cite{Schnitzer1993} \\
        \hline
            1993 & Thin-film chip architecture & Absorption & No & I, \cite{Schnitzer1993} \\
        \hline
            1996 & Patterned sapphire substrate  & Total reflection and absorption & Yes & I, \cite{Tadatomo2001} \\
        \hline
            Ongoing & Chip design for high LEE & Total reflection and absorption & No & I, \cite{Haerle2004} \\
        \hline
            $\sim$2000 & Silver p-contacts & Absorption & No & I, \cite{kondoh2001nitride} \\
        \hline
        \multicolumn{5}{c}{Internal quantum efficiency (IQE) $\eta_{IQ}$} \\
        \hline
            \textit{Year} & \textit{Innovation} & \textit{Area of improvement} & \textit{Spillover} & \textit{Source} \\
        \hline
            1994 & Double heterostructure & Higher radiative recombination probability & No & \cite{Nakamura1994} \\
        \hline
            1996 & Multiple quantum well & Higher radiative recombination probability & No & \cite{Koike1996} \\
        \hline
            Ongoing & Active region \newline Doping & Radiative recombination probability & No & I, \cite{schubert2018light} \\
        \hline
            Ongoing & Epitaxy \newline Improvements & Radiative recombination probability & No & I \\
        \hline
            Ongoing & Chip architecture \newline Improvements & Radiative recombination probability & No & I \\
        \hline
        \end{NiceTabularX}
        \vspace{2mm}
    \caption*{Note: \textit{Year} column indicates the first instance of application of corresponding invention in white LEDs. 'Ongoing' indicates improvements that are incremental in nature and have been ongoing since the earliest days of LED manufacturing, with no individual breakthroughs identified. \textit{Spillover} column indicates if the innovation involved technology spillovers. \textit{Source} column indicates the source of information about the innovation or improvement, with 'I' indicating expert interviews as such a source.}
    \vspace{-20mm}
    \label{tab:innovations_1}
\end{table}

\begin{table}[H]
    \caption{\textbf{LED innovations and technology improvements affecting key device sub-efficiencies.} Note that this table continues \cref{tab:innovations_1}.}
    \begin{NiceTabularX}{\textwidth}{ |l|X|X|l|l| }
    \hline
    \multicolumn{5}{c}{Droop $\eta_{droop}$} \\
        \hline
            ongoing & Lower current Density & Current density & No & I \\
        \hline
            ongoing & Epitaxy \newline Improvements & $\ominus/\oplus$ Distribution \newline Defect density & No & I, \cite{bhardwaj2016progress} \\
        \hline
            ongoing & Chip Architecture \newline Improvements  & $\ominus/\oplus$ Distribution \newline Defect density & No & I, \cite{Wildeson2017} \\
        \hline
            $<$ 2017 & Defect getting \newline underlayer & Defect density & No & I, \cite{haller2017burying} \\
        \hline
        \multicolumn{5}{c}{Light conversion efficiency (CE) $\eta_{C}$} \\
        \hline
            \textit{Year} & \textit{Innovation} & \textit{Area of improvement} & \textit{Spillover} & \textit{Source} \\
        \hline
            Ongoing & Lower current density & Current density & No & I \\
        \hline
            Ongoing & Epitaxy \newline Improvements & Charge distribution \newline Defect density & No & I, \cite{bhardwaj2016progress} \\
        \hline
            Ongoing & Chip architecture \newline improvements  & Charge distribution \newline Defect density & No & I, \cite{Wildeson2017} \\
        \hline
            $<$ 2017 & Defect gettering \newline underlayer & Defect density & No & I, \cite{haller2017burying} \\
        \hline
        \end{NiceTabularX}
        \vspace{2mm}
    \caption*{Note: \textit{Year} column indicates the first instance of application of corresponding invention in white LEDs. 'Ongoing' indicates improvements that are incremental in nature and have been ongoing since the earliest days of LED manufacturing, with no individual breakthroughs identified. \textit{Spillover} column indicates if the innovation involved technology spillovers. \textit{Source} column indicates the source of information about the innovation or improvement, with 'I' indicating expert interviews as such a source.}
    \label{tab:innovations_2}
\end{table}

@@

This section provides additional details for phosphor-related LED innovations identified in this study as affecting both spectral sub-efficiencies and consumer experience metrics. These include a detailed description of the history of those innovations, which we use to identify the details of involved technology spillovers. In addition, spectral data for the first identified LED products that used those innovations is presented in \cref{fig:phosphor_spectrum}, which we use to calculate consumer experience metrics, as well as LED spectral efficiency.

@@

Further efforts in LED phosphor development were directed towards synthesizing a red narrow-band phosphor. Narrow LED emission peak widths yield the highest luminous efficacy of radiation, as in this case less light is emitted in the far-red range of the spectrum in which the human eye is not very sensitive. After synthesizing several narrow-band phosphors emitting in yellow \cite{Hppe2004} and cyan \cite{Kechele2009}, the Schnick’s group identified the local cubic cation coordination structure of the cyan phosphor compound as the reason for its narrow band width \cite{lumi2016narrow}. A search for a structurally analogous nitride compound with the narrow red instead of the cyan emission was undertaken. After several unsuccessful attempts, the sought-after cuboidal nitride compound was found in a 2008 publication led by Francis DiSalvo \cite{Park2008Sr}. Based on information provided in this work, Schnick and colleagues synthesized and studied the spectral properties of a new narrow band red SLA phosphor in 2013 \cite{schmidt2013new}\cite{Pust2014}\cite{schmidt2017phosphors}. The material was introduced in commercial LED devices by Lumileds in 2015 \cite{lumi2016narrow_whitepaper}.

@@

\subsection{Metrics for Quantifying Technological Progress in LEDs}

@@

\textbf{Distinction between Metrics}

We use a well-established distinction between metrics. Those metrics which are related to the spectrum of the light source only ("color quality metrics" \cite{DLC_SSL_Requirements}\cite{dilaura2011lighting}), we group under the term consumer experience metrics. Those metrics, which are determined by the number and energy of photons leaving the device (as a function of the number of electrons entering it), we group under the term device performance metrics. As such, our distinction is physical at the fundamental level. This follows existing literature on the technological progress in solid-state lighting \cite{tsao2010solid}. We describe these metrics in detail in \cref{sec:device_performance_metrics}-\cref{sec:metrics_consumer}.

@@

A highly cited and frequently called-upon metric is \textit{flux/package}, as described in what has become known as Haitz's law. An observation rather than a law of nature, it describes increasing device performance and decreasing manufacturing cost. Originally introduced in 2000 \cite{haitz1999case}, it has been updated frequently since then. Unfortunately, the way Haitz's law is frequently used is  problematic due to the misleading conclusions it suggests. While the original version presented by Haitz displays the increasing performance of light-emitting diode chips, the graphs have been extended with datapoints from lamps which contain an unspecified number of chips. The metric therefore loses its value in describing technological progress at the chip-level. The continuing rise in performance inferred from this chart is disconnected from the actual, much lower, performance of single chips \cite{weinold2021compound}. This is shown in \cref{fig:haitz-org}-\cref{fig:haitz}.

\begin{figure}[h!]
\vspace{-10mm}
    \includegraphics[width=0.9\textwidth]{./figures/haitz_law_old_both.pdf}
    \caption{\textit{"Haitz's Law"}, as it is often reproduced in literature. The original data by Haitz for red LEDs from 2000 is indicated for reference. Note the different range on the abscissa compared to \cref{fig:haitz_new}. Sources: \cite{haitz1999case}\cite{haitz2011solid} and  \cref{fig:haitz_new}.}
    \label{fig:haitz-org}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.85\textwidth]{./figures/haitz_law_white.pdf}
    \caption{Adaption of \textit{"Haitz's Law"}. Note that the increase in flux per single chip is not increasing as starkly as the flux for multi-chip packages. Shown are datapoints for best commercial performers. See \cref{sec:sources} for the full list of sources and references.}
    \label{fig:haitz_new}
\end{figure}

@@

During the years following the market introduction of white LEDs, the primary metric of progress in solid-state lighting was typically luminous flux. This was because the luminous flux (total brightness) of early white LEDs was too small to allow for the economical combination of multiple LEDs into lamps for general illumination purposes. In 2000, the highest performing devices yielded around 10 lm, just below the output of a candle as defined in the unit candela (1 cd=12.57 lm) \cite{haitz2011solid}. Progress in this metric was commonly rendered together with the associated reduction in retail prices as “Haitz’s Law”\cite{haitz1999case,haitz2011solid}. Today, however, the devices with the highest performance yield in excess of 1600 lm, the equivalent of a 100 W incandescent bulb \cite{cree2020bright}. LED brightness has thereby become sufficient to enable the construction of lamps from multiple LED devices, with contemporary improvements focusing instead on higher efficiency and quality of light instead of brightness. Even though a large number of scientific publications and industry periodicals continue to focus on brightness as a metric of progress in lighting, we find that, at this point in time, this metric is insufficient to capture the complexity of the multitude of efficiency improvements that have been driving the overall LED efficiency \cite{weinold2021compound}. Specifically, the stagnating levels of luminous flux in the highest performing devices do not capture the outcomes of a major area of LED research, which is improving electrical efficiency at constant brightness.

We therefore use a combination of the total device efficiency ("lamp efficiency") and the sub-efficiencies that describe different physical loss channels within the device to describe progress in LED technology. In line with previous publications \cite{schubert2018light,tsao2010solid}, we use the sub-efficiencies: forward voltage efficiency $\eta_{V_f}$, light extraction efficiency $\eta_{LE}$, internal quantum efficiency $\eta_{IQ}$, droop $\eta_{droop}$, conversion efficiency $\eta_{C}$, spectral efficiency $\eta_{S}$. The equations defining these sub-efficiencies are provided below. The overall efficiency ("Lamp Efficiency") $\eta_L$ of a light-emitting diode package is the product of all considered sub-efficiencies:
%
\begin{equation}
    \eta_L = \prod_{i=(V_f, \dots, S)} \eta_i
\end{equation}
%
This metric describes the cumulative electrical and optical losses within the device, as well as the light conversion losses in the phosphor layer.

@@

Internal quantum efficiency (IQE) describes non-radiative recombinatory processes in the semiconductor bulk. It is defined as
%
\begin{equation}
    \eta_{IQ} = \frac{\# \text{of photons created/s}}{\# \text{of electron-hole pairs in/s}}
\end{equation}

@@

Light conversion efficiency (LCE) describes the sum of all losses in the process of converting the blue light that is the basis for all phosphor-converted white LEDs. Specifically, energy is lost due to three distinct effects. The first effect is the inherent wavelength-conversion loss or \textit{Stokes shift} as the difference between the original blue wavelength and the down-converted red, yellow or green wavelength. The second effect are non-radiative processes in the phosphor material. The third , scattering and absorption processes in the phosphor material. These three effects are listed individually in Figure 4. The overall light conversion efficiency is defined as
%
\begin{equation}
    \eta_{C} = \frac{E_{\textcolor{blue}{B}}}{\sum_{i=\textcolor{red}{R},\textcolor{orange}{O},\textcolor[RGB]{225, 180, 0}{Y},\textcolor{teal}{G}} E_i}
\end{equation}
%
where $E_{\textcolor{blue}{B}}$ is the total energy of blue light before down-conversion and $E_i$ denotes the total energy of light at color $i$ corresponding to the down-converted photon wavelength \cite{schubert2018light}\cite{tsao2010solid}. Since every individual phosphor component in the device has its own associated conversion losses, the denominator sums over all phosphor components, i.e. \textcolor{red}{R}ed, \textcolor{orange}{O}range, \textcolor[RGB]{225, 180, 0}{Y}ellow, \textcolor{teal}{G}reen.

@@

We address the limitations of data availability for the chip-level LED manufacturing costs and sales prices by developing and applying a bottom-up LED manufacturing cost model with process-step resolution. In our case, it is enabled by the general availability of historical data on prices of relevant raw materials, components, and manufacturing equipment, and the relative cost data on manufacturing processes, which is occasionally published as part of industry press releases \cite{ledinside2013csp}\cite{seoul2015csp}.

We describe our manufacturing cost model in the Methods section in the main text of the article, and discuss its details, including its structure and equations, manufacturing process flows for the chip architectures under consideration, input data, explanation of our cost modelling approach based on yielded costs, as well as the model’s limitations, in \cref{sec:costmodel_details}. The outcome of the cost model is the cumulative yielded manufacturing cost per LED package, which includes all costs associated with producing the chip, including running costs of the factory.

@@

\textit{Semi}, the global industry association representing the electronics and semiconductor manufacturing industry, has published standards for manufacturing cost modeling. These include term definitions and equations for calculating cost and the associated quantities, such as equipment depreciation. This standard has been widely adopted industry and provides a standardized comparative analysis of the economics of the manufacturing chain \cite{ragona2002cost}\cite{wouters2018industry}. An overview of the parameters used in the model, according to \textit{SEMI Standard 35} is provided in \cref{semi_pyramid}. Note that the full ensemble of inputs are required only for an industry cost model used by manufacturers for detailed forecasting. As per the scope of our model, we need only a selection of these inputs. Any inputs unrelated to the manufacturing process itself can be disregarded. These include taxes, equipment installation, delivery facilities, customer orders etc.

While generally comprehensive, the model is limited by the availability of performance data for the most recent manufacturing equipment, as well as the confidential nature of cost structures of Asia-based manufacturing lines. A complete treatment of the limitations of the model is provided in \cref{sec:costmodel_limitations}.

>>>>

The historical development development of light-emitting diodes from a novelty semiconductor experiment into powerful lighting devices has received a considerable amount of attention following the recognition of pioneering work on blue LED by Japanese researchers Shuji Nakamura, Isamu Akasaki and Hiroshi Amano with the 2014 Nobel Prize in Physics \cite{Nakamura2015}. The sources and disaggregated contributions of subsequent improvements in white LED technology, however, have not been documented in the literature in a systematic fashion, though there were notable publications reporting on the progress in the design of devices \cite{Shchekin2006,krames2007led,laubsch2009high,hahn2014development}, or the technological improvements  underlying the improvement in overall device efficiency for best performers in 2009 \cite{tsao2010solid} and 2016 \cite{pattison2017solid}.

@@

Due to these limitations, any recommendations on future improvements in white LED technology based on these studies are fragmented at best. Furthermore, the effect of individual innovations and technology spillovers, which has been investigated in solar PV \cite{kavlak2018evaluating,kolesnikov2021spillovers,nemet2019solar}, and to some extent in lithium-ion batteries \cite{Stephan2021}, has not yet been studied in the context of lighting.

In our study, we overcome the limitations of previous literature by, first, synthesizing data provided in multiple publications across an ensemble of various cost, performance and consumer experience metrics for tracking progress in white LED technology over an extended period of time. Second, our key advantage and contribution compared to these publications is a multi-method approach that we take to fill in the data gaps and augment information obtained from the literature sources by our own performance calculations, a bottom-up cost model and a series of expert interviews. This enabled us to synthesize evidence from multiple sources and, for the first time, reconstruct a comprehensive picture of the historical progress in white LED technology throughout its entire history.

@@

\begin{figure}[H]
\centering
\includegraphics[width=15.5cm]{./figures/subefficiencies_progress.pdf}
\caption{\textbf{Historical improvements in white LED sub-efficiencies.} Panels A1-A2 show data for key metrics used to calculate forward voltage efficiency $\eta_{Vf}$ and droop efficiency $\eta_{droop}$. Panels B1-B2 show data for internal quantum efficiency $\eta_{IQ}$ and light extraction efficiency $\eta_{LE}$. Spectral data related to spectral sub-efficiency is shown in \cref{fig:phosphor_spectrum}. Additional context for the panels is provided below. Source: own synthesis of data from the list of sources provided in \cref{sec:sources}. For an illustration of the different chip architectures shown, compare \cref{sec:chip_architecture}. Data on average device performance adapted from U.S. Department of Energy (DOE) Reports \cite{doe_ssl_multiyear_2007}\cite{doe_ssl_multiyear_2008}\cite{doe_ssl_multiyear_2013}\cite{doe_ssl_rnd_2016}\cite{doe_ssl_rnd_2018} and an Osram company report \cite{beale_leds_2015}.}
\label{fgr:subeff}
\vspace{-20mm}
\end{figure}

@@

\begin{table}[H]
    \caption{\textbf{White LED innovations and technology improvements affecting key device sub-efficiencies.} Note this table is continued in \cref{tab:innovations_2}.}
    \begin{NiceTabularX}{\textwidth}{ |l|X|X|l|l| }
        \hline
        \multicolumn{5}{c}{Forward Voltage Efficiency (FVE) $\eta_{Vf}$} \\
        \hline
            \textit{Year} & \textit{Innovation} & \textit{Area of Improvement} & \textit{Spillover} & \textit{Source} \\
        \hline
            1999 & Indium Tin Oxide \newline Current spreading layer & Contact resistance & Yes & \cite{margalith1999indium}\\
        \hline
            1998 & Digitated electrodes & Contact resistance & No & \cite{steigerwald2001electrode} \\
        \hline
            Ongoing & Epitaxy improvements \newline and better doping & Polarization and \newline bulk resistance & No & I \\
        \hline
            1998 & Silver p-Contacts & Contact resistance & No & \cite{kondoh2001nitride} \\
        \hline
        \multicolumn{5}{c}{Light extraction efficiency (LEE) $\eta_{LE}$} \\
        \hline
            \textit{Year} & \textit{Innovation} & \textit{Area of improvement} & \textit{Spillover} & \textit{Source} \\
        \hline
            $<$ 2003 & Optimization for \newline cavity effects & Reduces self-interference of quantum well & No & I, \cite{Shen2003} \\
        \hline
            1993 & Chip surface randomization & Total reflection and absorption & No & I, \cite{bergh1973surface}\cite{Schnitzer1993} \\
        \hline
            1993 & Thin-film chip architecture & Absorption & No & I, \cite{Schnitzer1993} \\
        \hline
            1996 & Patterned sapphire substrate  & Total reflection and absorption & Yes & I, \cite{Tadatomo2001} \\
        \hline
            Ongoing & Chip design for high LEE & Total reflection and absorption & No & I, \cite{Haerle2004} \\
        \hline
            $\sim$2000 & Silver p-contacts & Absorption & No & I, \cite{kondoh2001nitride} \\
        \hline
        \multicolumn{5}{c}{Internal quantum efficiency (IQE) $\eta_{IQ}$} \\
        \hline
            \textit{Year} & \textit{Innovation} & \textit{Area of improvement} & \textit{Spillover} & \textit{Source} \\
        \hline
            1994 & Double heterostructure & Higher radiative recombination probability & No & \cite{Nakamura1994} \\
        \hline
            1996 & Multiple quantum well & Higher radiative recombination probability & No & \cite{Koike1996} \\
        \hline
            Ongoing & Active region \newline Doping & Radiative recombination probability & No & I, \cite{schubert2018light} \\
        \hline
            Ongoing & Epitaxy \newline Improvements & Radiative recombination probability & No & I \\
        \hline
            Ongoing & Chip architecture \newline Improvements & Radiative recombination probability & No & I \\
        \hline
        \end{NiceTabularX}
        \vspace{2mm}
    \caption*{Note: \textit{Year} column indicates the first instance of application of corresponding invention in white LEDs. 'Ongoing' indicates improvements that are incremental in nature and have been ongoing since the earliest days of LED manufacturing, with no individual breakthroughs identified. \textit{Spillover} column indicates if the innovation involved technology spillovers. \textit{Source} column indicates the source of information about the innovation or improvement, with 'I' indicating expert interviews as such a source.}
    \vspace{-20mm}
    \label{tab:innovations_1}
\end{table}

\begin{table}[H]
    \caption{\textbf{White LED innovations and technology improvements affecting key device sub-efficiencies.} Note that this table continues \cref{tab:innovations_1}.}
    \begin{NiceTabularX}{\textwidth}{ |l|X|X|l|l| }
    \hline
    \multicolumn{5}{c}{Droop $\eta_{droop}$} \\
        \hline
            ongoing & Lower current Density & Current density & No & I \\
        \hline
            ongoing & Epitaxy \newline Improvements & $\ominus/\oplus$ Distribution \newline Defect density & No & I, \cite{bhardwaj2016progress} \\
        \hline
            ongoing & Chip Architecture \newline Improvements  & $\ominus/\oplus$ Distribution \newline Defect density & No & I, \cite{Wildeson2017} \\
        \hline
            $<$ 2017 & Defect getting \newline underlayer & Defect density & No & I, \cite{haller2017burying} \\
        \hline
        \multicolumn{5}{c}{Light conversion efficiency (CE) $\eta_{C}$} \\
        \hline
            \textit{Year} & \textit{Innovation} & \textit{Area of improvement} & \textit{Spillover} & \textit{Source} \\
        \hline
            Ongoing & Lower current density & Current density & No & I \\
        \hline
            Ongoing & Epitaxy \newline Improvements & Charge distribution \newline Defect density & No & I, \cite{bhardwaj2016progress} \\
        \hline
            Ongoing & Chip architecture \newline improvements  & Charge distribution \newline Defect density & No & I, \cite{Wildeson2017} \\
        \hline
            $<$ 2017 & Defect gettering \newline underlayer & Defect density & No & I, \cite{haller2017burying} \\
        \hline
        \end{NiceTabularX}
        \vspace{2mm}
    \caption*{Note: \textit{Year} column indicates the first instance of application of corresponding invention in white LEDs. 'Ongoing' indicates improvements that are incremental in nature and have been ongoing since the earliest days of LED manufacturing, with no individual breakthroughs identified. \textit{Spillover} column indicates if the innovation involved technology spillovers. \textit{Source} column indicates the source of information about the innovation or improvement, with 'I' indicating expert interviews as such a source.}
    \label{tab:innovations_2}
\end{table}

@@

This section provides additional details for phosphor-related LED innovations identified in this study as affecting both spectral sub-efficiencies and consumer experience metrics. These include a detailed description of the history of those innovations, which we use to identify the details of involved technology spillovers. In addition, spectral data for the first identified white LED products that used those innovations is presented in \cref{fig:phosphor_spectrum}, which we use to calculate consumer experience metrics, as well as LED spectral efficiency.

@@

Further efforts in LED phosphor development were directed towards synthesizing a red narrow-band phosphor. Narrow LED emission peak widths yield the highest luminous efficacy of radiation, as in this case less light is emitted in the far-red range of the spectrum in which the human eye is not very sensitive. After synthesizing several narrow-band phosphors emitting in yellow \cite{Hppe2004} and cyan \cite{Kechele2009}, the Schnick’s group identified the local cubic cation coordination structure of the cyan phosphor compound as the reason for its narrow band width \cite{lumi2016narrow}. A search for a structurally analogous nitride compound with the narrow red instead of the cyan emission was undertaken. After several unsuccessful attempts, the sought-after cuboidal nitride compound was found in a 2008 publication led by Francis DiSalvo \cite{Park2008Sr}. Based on information provided in this work, Schnick and colleagues synthesized and studied the spectral properties of a new narrow band red SLA phosphor in 2013 \cite{schmidt2013new}\cite{Pust2014}. The material was introduced in commercial white LED devices by Lumileds in 2015 \cite{lumi2016narrow_whitepaper}.

@@

\subsection{Metrics for Quantifying Technological Progress in LEDs}
\label{sec:metrics}

@@

\textbf{Distinction between Metrics of Efficiency and Consumer Experience}

We use a well-established distinction between the two types of LED lighting technology performance metrics. Those metrics which are related to the spectrum of the light source only ("color quality metrics" \cite{DLC_SSL_Requirements}\cite{dilaura2011lighting}), we group under the term consumer experience metrics. Those metrics, which are determined by the number and energy of photons leaving the device (as a function of the number of electrons entering it), we group under the term LED efficiency metrics. As such, our distinction is physical at the fundamental level. This follows existing literature on the technological progress in solid-state lighting \cite{tsao2010solid}. We describe these metrics in detail in \cref{sec:device_performance_metrics}-\cref{sec:metrics_consumer}.

@@

A highly cited and frequently called-upon metric is \textit{flux per package}, as described in what has become known as Haitz's law. An observation rather than a law of nature, it describes increasing device performance and decreasing manufacturing cost. Originally introduced in 2000 \cite{haitz1999case}, it has been updated frequently since then. Unfortunately, the way Haitz's law is frequently used is  problematic due to the misleading conclusions it suggests. While the original version presented by Haitz displays the increasing performance of light-emitting diode chips, the graphs have been extended with datapoints from lamps which contain an unspecified number of chips. The metric therefore loses its value in describing technological progress at the chip-level. The continuing rise in performance inferred from this chart is disconnected from the actual, much lower, performance of single chips \cite{weinold2021compound}. This is shown in \cref{fig:haitz-org}-\cref{fig:haitz_new}.

\begin{figure}[h!]
\vspace{-10mm}
    \includegraphics[width=0.9\textwidth]{./figures/haitz_law_old_both.pdf}
    \caption{\textit{"Haitz's Law"}, as it is often reproduced in literature. The original data by Haitz for red LEDs from 2000 is indicated for reference. Note the different range on the horixontal axis compared to \cref{fig:haitz_new}. Sources: \cite{haitz1999case}\cite{haitz2011solid} and  \cref{fig:haitz_new}.}
    \label{fig:haitz-org}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.85\textwidth]{./figures/haitz_law_white.pdf}
    \caption{Corrected version of \textit{"Haitz's Law"}. Note that the increase in flux per single chip is not increasing as starkly as the flux for multi-chip packages. Shown are datapoints for best commercial performers. See \cref{sec:sources} for the full list of sources and references.}
    \label{fig:haitz_new}
\end{figure}

@@

During the years following the market introduction of white LEDs, the primary metric of progress in solid-state lighting was typically luminous flux. This was because the luminous flux (total brightness) of early white LEDs was too small to allow for the economical combination of multiple LEDs into lamps for general illumination purposes. In 2000, the highest performing devices yielded around 10 lm, just below the output of a candle as defined in the unit candela (1 cd=12.57 lm) \cite{haitz2011solid}. Progress in this metric was commonly rendered together with the associated reduction in retail prices as “Haitz’s Law”\cite{haitz1999case,haitz2011solid}, which we discuss in \cref{sec:metrics}. Today, however, the devices with the highest performance yield in excess of 1600 lm, the equivalent of a 100 W incandescent bulb \cite{cree2020bright}. LED brightness has thereby become sufficient to enable the construction of lamps from multiple LED devices, with contemporary improvements focusing instead on higher efficiency and quality of light instead of brightness. Even though a large number of scientific publications and industry periodicals continue to focus on brightness as a metric of progress in lighting, we find that, at this point in time, this metric is insufficient to capture the complexity of the multitude of efficiency improvements that have been driving the overall LED efficiency \cite{weinold2021compound}. Specifically, the stagnating levels of luminous flux in the highest performing devices do not capture the outcomes of a major area of LED research, which is improving electrical efficiency at constant brightness.

We therefore use a combination of the total device efficiency ("lamp efficiency") and the sub-efficiencies that describe different physical loss channels within the device to describe progress in LED technology. In line with previous publications \cite{schubert2018light,tsao2010solid}, we use the following sub-efficiencies: forward voltage efficiency $\eta_{V_f}$, light extraction efficiency $\eta_{LE}$, internal quantum efficiency $\eta_{IQ}$, droop $\eta_{droop}$, conversion efficiency $\eta_{C}$, spectral efficiency $\eta_{S}$. The equations defining these sub-efficiencies are provided below. The overall efficiency ("Lamp Efficiency") $\eta_L$ of a light-emitting diode package is the product of all considered sub-efficiencies:
%
\begin{equation}
    \eta_L = \prod_{i=(V_f, \dots, S)} \eta_i
\end{equation}
%
This metric describes the cumulative electrical and optical losses within the device, as well as the light conversion losses in the phosphor layer.

@@

Internal quantum efficiency (IQE) describes non-radiative recombinatory processes in the semiconductor bulk. It is defined as
%
\begin{equation}
    \eta_{IQ} = \frac{\# \text{of photons created/s}}{\# \text{of electron-hole pairs injected/s}}
\end{equation}

@@

Light conversion efficiency (LCE) describes the sum of all losses in the process of converting the blue light that is the basis for all phosphor-converted white LEDs. Specifically, energy is lost due to three distinct effects. The first effect is the inherent wavelength-conversion loss or \textit{Stokes shift} as the difference between the original blue wavelength and the down-converted red, yellow or green wavelength. The second effect are non-radiative processes in the phosphor material. The third , scattering and absorption processes in the phosphor material. These three effects are listed individually in Figure 4 in the main text. The overall light conversion efficiency is defined as
%
\begin{equation}
    \eta_{C} = \frac{E_{\textcolor{blue}{B}}}{\sum_{i=\textcolor{red}{R},\textcolor{orange}{O},\textcolor[RGB]{225, 180, 0}{Y},\textcolor{teal}{G}} E_i}
\end{equation}
%
where $E_{\textcolor{blue}{B}}$ is the total energy of blue light before down-conversion and $E_i$ denotes the total energy of light at color $i$ corresponding to the down-converted photon wavelength \cite{schubert2018light}\cite{tsao2010solid}. Since every individual phosphor component in the device has its own associated conversion losses, the denominator sums over all phosphor components, i.e. \textcolor{red}{R}ed, \textcolor{orange}{O}range, \textcolor[RGB]{225, 180, 0}{Y}ellow, \textcolor{teal}{G}reen.

@@

We address the limitations of data availability for the chip-level LED manufacturing costs and sales prices by developing and applying a bottom-up white LED manufacturing cost model with process-step resolution. In our case, it is enabled by the general availability of historical data on prices of relevant raw materials, components, and manufacturing equipment, and the relative cost data on manufacturing processes, which is occasionally published as part of industry press releases \cite{ledinside2013csp}\cite{seoul2015csp}.

We describe our manufacturing cost model in the Methods section in the main text of the article, and discuss its details, including its structure and equations, manufacturing process flows for the chip architectures under consideration, input data, explanation of our cost modelling approach based on yielded costs, as well as the model’s limitations, in \cref{sec:costmodel_structure}. The outcome of the cost model is the cumulative yielded manufacturing cost per LED package, which includes all costs associated with producing the chip, including running costs of the factory.

@@

\textit{Semi}, the global industry association representing the electronics and semiconductor manufacturing industry, has published standards for manufacturing cost modeling. These include term definitions and equations for calculating cost and the associated quantities, such as equipment depreciation. This standard has been widely adopted industry and provides a standardized comparative analysis of the economics of the manufacturing chain \cite{ragona2002cost}\cite{wouters2018industry}. An overview of the parameters used in the model, according to \textit{SEMI Standard 35} is provided in \cref{fig:semi_pyramid}. Note that the full ensemble of inputs are required only for an industry cost model used by manufacturers for detailed forecasting. As per the scope of our model, we need only a selection of these inputs. Any inputs unrelated to the manufacturing process itself can be disregarded. These include taxes, equipment installation, delivery facilities, customer orders etc.

While generally comprehensive, the model is limited by the availability of performance data for the most recent manufacturing equipment, as well as the confidential nature of cost structures of Asia-based manufacturing lines. A complete treatment of the limitations of the model is provided in \cref{sec:costmodel_structure}.
